\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{float}
\usepackage{caption}
\usepackage{titlesec}
\usepackage{fancyhdr}
\usepackage{amsmath}
\usepackage{pgfplots}

\hypersetup{colorlinks=true,linkcolor=black,citecolor=black,urlcolor=blue}
\setstretch{1.15}
\setlength{\parskip}{0.5em}
\setlength{\parindent}{0pt}
\titleformat{\section}{\large\bfseries}{\thesection}{0.75em}{}
\titleformat{\subsection}{\normalsize\bfseries}{\thesubsection}{0.75em}{}
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{QNX Quantum Secure OTA Update}
\fancyhead[L]{Comp 4900E}
\fancyfoot[C]{\thepage}

\begin{document}

\begin{titlepage}
    \centering
    {\Large Carleton University \par}
    \vspace{1.5cm}
    {\LARGE \textbf{QNX Quantum Secure OTA Update} \par}
    \vspace{0.5cm}
    {\large COMP4900E, Real Time Operating Systems \par}
    \vspace{1.25cm}
    {\large Final Project Report\par}
    \vspace{1.5cm}
    \begin{tabular}{ll}
        Ben Whitten & \texttt{BenWhitten@cmail.carleton.ca} \\
        Jacob Bonner & \texttt{JacobBonner@cmail.carleton.ca} \\
        Caitlin Wardle & \texttt{CaitlinWardle@cmail.carleton.ca} \\
    \end{tabular}
    \vfill
    {\large \today \par}
\end{titlepage}
\begin{abstract}
    This project investigates whether a post-quantum digital signature scheme can be integrated into a QNX over-the-air (OTA) update workflow without making the update process impractical for embedded devices. Our target platform is a Raspberry Pi 4 running QNX OS 8.0, and our core comparison is between three update modes, unsigned OTA, classically signed OTA using RSA and Post Quantum Computing OTA using ML-DSA. The system is designed around an A/B partition update model in which a device checks a remote server for a new image, downloads an update bundle, verifies its signature, writes the new image to the inactive partition and reboots into the updated system if the validation succeeds. The report presents the need for secure OTA, the quantum risks facing current signing techniques, the benefits of ML-DSA, the implementation architecture and the comparisons across algorithms. Our evaluation focuses on signature generation time, signature verification time, total OTA update time, CPU usage, RAM usage, signature size, and storage overhead.
\end{abstract}
\tableofcontents
\newpage
\section{Introduction}
Embedded systems in automotive, medical, industrial and IoT environments usually rely on remote software updates to remain secure and functional over their deployment lifespan. Over the air (OTA) updates allow new software images or configuration changes to be delivered over a network, so physical access of the device is not required. OTA typically involves packaging an update on a server, transmitting it to a device, verifying its integrity and authenticity on the device, installing it and finally rebooting into the new version. For embedded systems OTA is beneficial because it reduces maintenance cost, minimizes system downtime, allows for rapid deployment and satisfies consumer need for at home maintenance. However, OTA has one critical problem, if an attacker can forger or tamper with an update, then they can gain persistent control of the device, a potentially disastrous scenario. This risk is exacerbated by the long lifespans of many embedded devices, since cryptographic assumptions used today may no longer be secure over the full life of the product. Most OTA sustems rely on classical public key encryption schemes, like RSA or elliptic-curve cryptography. But the long term security of these algorithms is threatened by large scale quantum computing, since new methods, like Shor's algorithm would break the hard mathematical problems these encryption schemes currently depend on. As such it is important to investigate whether post quantum signature mechanisms can be integrated into OTA validation, without unacceptable cost to the devices performance. This project aims to address that question by designing and evaluating a secure OTA mechanism for QNX based embedded systems using a Post Quantum Computing (PQC) signature algorithm for update validation. Our implementation is for QNX OS 8.0 on a Rasberry Pi 4, and uses ML-DSA through liboqs as the pqc signature scheme, with RSA through OpenSSL as the classical comparison baseline. As a baseline we have also included unsigned OTA, to isolate the cost of signature validation from the cost of the update process. 
\subsection{Research Question}
Is it feasible to replace a classical code signing algorithm in a QNX OTA update with a post quantum computing signature scheme without causing significant performance degradation on a simple embedded device?
\subsection{Objectives}
\begin{enumerate}[leftmargin=2em]
    \item Implement a working OTA update prototype on QNX 8.0 for Raspberry Pi 4.
    \item Integrate a digital signature verification step into the OTA validation path.
    \item Implement RSA signing and verification as a baseline.
    \item Implement ML-DSA signing and verification using liboqs as the post quantum solution
    \item Compare unsigned, RSA signed and ML-DSA signed updates using timing and resource usage.
    \item Analyze the resource demands of post quantum computing signatures on OTA updates in embedded devices to determine a baseline of resource and time requirements. As individual embedded systems have specific requirements the praticality of these drains will need to be determined on a per system basis.  
\end{enumerate}
\subsection{Scope}
The scope of this project is the OTA image packaging, download, signature validation, installation to an inactive partition and the reboot on an embedded device. The focus is on code signing and verification, not on transport encryption, key exchange or large scale deployments. Within this scope the project has three parts. Firstly it builds an OTA pipeline for QNX on Raspberry Pi 4 using an A/B partition model. Secondly it builds both a classical (RSA) and a post quantum computing (ML-DSA) signature into the update validation mechanism. Finally, it collects data pursuant to algorithm cost and device level measurements like runtime, CPU load, memory consumption and storage overhead.  
\subsection{Contributions}
\begin{itemize}[leftmargin=2em]
    \item \textbf{Ben Whitten:} \textcolor{red}{fill in}
    \item \textbf{Jacob Bonner:} \textcolor{red}{fill in}
    \item \textbf{Caitlin Wardle:} \textcolor{red}{fill in}
\end{itemize}
\section{Background}
\subsection{OTA Updates in Embedded Systems}
OTA updates allow a device to receive a new software image over a network connection
rather than through physical access. A typical OTA pipeline packages an image on
a server, transfers the image to the embedded device, validates the update on the device, installs the update, and reboots into the new version. For embedded systems,
OTA is especially useful for remote or critical devices. 
Despite the benefits, OTA updates are more vulnerable to installation failures and malicious attacks. If a device loses power, writes an incomplete
image, or boots an invalid update, it may become unusable or “bricked.” For this reason,
OTA systems rely on rollback capabilities,saved versions, and cryptographic
verification. A secure OTA design must protect against accidental
failure and against malicious changes. 
\subsection{Why Quantum Resistance Matters}
The signature verification step in current OTA systems uses classic public key
cryptography. The device stores a trusted public key and uses it to verify that the received
update was signed by the trusted private key holder. This model works well today, but its
long term security depends on mathematical formulas that future quantum computers may be able to break.
Shor’s algorithm would make integer factorization and elliptic curve discrete
logarithms solvable, destroying the security of RSA and ECC.
Since many embedded systems have long life spans, a device currently in use may still rely on these algorithms long after quantum attacks become effective. The \textit{harvest now, forge later}
threat model proposes attackers can collect update data and system images now and break them in the future, when quantum computers are developed. \textcolor{red}{Continue/make better}
\subsection{Post Quantum Signatures}
Post quantum computing (PQC) refers to cryptographic algorithms designed to remain secure even against attackers with quantum computers. For digital signatures the major categories of PQC are lattice based, hash based and multivariate constructions. This project focuses on ML-DSA, a lattice based digital signature algorithm  standardized by NIST and supported 
by liboqs.
ML-DSA was chosen because it offers a strong balance between security and practical performance for constrained systems. Compared to other algorithms like SLH-DSA, ML-DSA
generally provides smaller signatures and faster signing and verification, which are important when working with embedded devices operating with limited resources.
\subsection{Classical Signatures}
To determine the baseline overhead of current validation mechanisms this project has also implemented a classical scheme (RSA) that reflects present day OTA validations. The project uses RSA via OpenSSL as our baseline  baseline classical signature mechanism. RSA is well understood, widely supported, and easy to implement in a comparable signing and verification workflow. 
\section{System Design}
\subsection{Hardware}
The implementation was preformed with a Rasberry PI 4 running QNX OS 8.0. The Raspberry Pi 4 was chosen because it is affordable, well documented and practical for testing simple embedded devices. Using real hardware allows for device level measurements to realistically capture the performance limits and operational constrains of actual embedded systems.  
\subsection{A/B Partition OTA Model}
The OTA system is designed around an A/B partition architecture. In this model,
partition A contains the currently running system image, while partition B is inactive and can
be used as the destination for a downloaded update. Firstly the device will check the remote server for a new image. If a new image is available, the device downloads the update package. Then the device verifies the digital signature associated with the update. If the verification is successful, the new image is written to inactive partition B. The boot configuration is changed so the device boots from the updated partition B. The system reboots into the new image. This design has excellent rollback capabilities, if the verification fails, installation is aborted. If the update is written but becomes unstable, the system can fall back to the previously working partition. This makes the A/B model robust, and ideal for OTA. 
\subsection{Threat Model}
\textcolor{red}{What can go wrong/how can malicious users exploit the system and how are we solving it}
\subsection{ Cryptographic Workflow}
The Cryptographic workflow consists of two sides, the server side packaging step and the device side verification step.
\textbf{Server Side:}
\begin{enumerate}[leftmargin=2em]
\item Build or package the QNX image using \texttt{mkqnximage}
\item Compute a SHA 256 digest of the image
\item Sign the digest using RSA or ML-DSA
\item Publish the update image, signature and metadata to the server
\end{enumerate}
\textbf{Device Side:}
\begin{enumerate}[leftmargin=2em]
\item Download the update package from server.
\item Recompute the SHA 256 digest of the image.
\item Verify the signature using the \textcolor{red}{stored} public key.
\item Abandon installation if verification fails.
\item If verification succeeds, write the image to partition B and reboot.
\end{enumerate}
\section{Implementation}
\subsection{Software}
The main technologies used in this project are QNX Momentics IDE, \texttt{mkqnximage}, liboqs, OpenSSL, Git/GitHub, and Raspberry Pi 4 hardware. QNX Momentics is used as the main development environment. \texttt{mkqnximage} is used to generate QNX operating system images. OpenSSL provides classical RSA signing and verification, while liboqs provides the ML-DSA implementation for post quantum computing validation. Github \textcolor{red}{host the remote server} and acts as storage and version control for the project implementation. 
\subsection{OTA Update}
The unsigned OTA serves as a baseline measurement. In this mode, the device checks for an update, downloads the image, and writes it to the inactive partition without performing any kind of validation. This will establish a minimum time and resource cost of the OTA process itself. Without this baseline this project cannot accurately determine the cost of update retrieval and image download from the cost of the update validation mechanism 
\subsection{RSA OTA Validation}
The RSA method continues the baseline OTA mechanism, but adds a signing step on the server and a verification step on the device. On the server a SHA 256 digest of the update image is signed with an RSA private key through OpenSSL. On the device, the update image is hashed again and the signature is verified  with the RSA public key before installation proceeds. This models the classical public key verification currently used in many secure update systems. It provides a comparison point for evaluating the increased cost of post quantum signatures. 
\subsection{ML-DSA OTA Validation}
The post quantum method follows the same logic as RSA but replaces the signature scheme with ML-DSA using liboqs. On the server, the image digest is signed using an ML-DSA private key. On the device, the update digest is recomputed and verified using the corresponding ML-DSA public key. This implementation aims to keep workflow as consistent as possible while only swapping the signature scheme, in order to get a clean comparison of algorithmic cost. The main difference between RSA and ML-DSA is signature size and key size and we suspect this to have a negative impact on verification time.
\subsection{Metrics}
The following metrics are collected for each update mode:
\begin{itemize}[leftmargin=2em]
    \item \textbf{Signature Generation Time}: Time required to create the signature on the server side.
    \item \textbf{Signature Verification Time}: Time required to verify the signature on the device. 
    \item \textbf{Total OTA Time}: Full Time required from checking for updates/downloading update through validation to installation.
    \item \textbf{CPU Usage}: CPU cost observed during validation.
     \item \textbf{RAM Usage}: Maximum memory consumed during validation.
     \item \textbf{Signature Size}: Size in bytes of the generated signature.
     \item \textbf{Storage Overhead}: Additional package storage needed for signature components and keys. 
\end{itemize}
\subsection{Measurement Procedure}
To test the impact of post quantum computing signature algorithms on QNX OTA updates, the update code will be run using each of the decided upon signature schemes: RSA, MLDSA, and unsigned. Each signature scheme will be run 20 times and the metrics listed above in section 4.5 will be collected across each test.
\subsection{Data Analysis Procedure}
Once the data has been collected according to what's outlined in section 4.6, the mean and standard error of each metric across all tests will be calculated. These results will then be used in sections 6 and 7 to determine whether the adoption of post-quantum cryptographic algorithms is currently feasible or not.
\section{OTA Tutorial}
\text{{The following are the requirements and steps wto replicate our OTA tests.}}
\subsection{Hardware \& Software Requirements}
\noindent
\textbf{Github Repo:} \url{https://github.com/BenjaminWhitten/COMP4900-Quantum-Secure-OTA} \\
\textbf{Host Server PC:} Windows 11 (with WSL Ubuntu) \\
\textbf{QNX SDP:} Version 8.0 (requires BlackBerry QNX license) \\
\textbf{QNX Raspberry Pi Board Support:} QNX SDP 8.0 BSP for Raspberry Pi BCM2711 R-PI4 \\
\textbf{Development IDE:} QNX Momentics IDE \\
\textbf{QNX OS Terminal Interface:} PuTTY \\
\textbf{Local Server:} Windows OpenSSH Server \\
\textbf{Micro Controller:} Raspberry Pi 4 \\
\textbf{Bootloader:} \url{https://github.com/u-boot/u-boot} \\
\textbf{Raspberry Pi Firmware:} \url{https://github.com/raspberrypi/firmware.git}

\noindent
\textbf{Required Cables/Accessories:}
\begin{itemize}
    \item USB to TTL serial cable (to access the Pi’s UART pins for console)
    \item Micro SD card (32 GB)
    \item Ethernet Cable
\end{itemize}
\subsection{Install Necessary Software Tools}
The first thing to do will be to make sure all the required software is downloaded.\\\\
\textbf{1) Installing QNX SDP 8.0 and BSP:}
\begin{enumerate}
    \item Make sure you have downloaded the QNX Software Center from the QNX website. To do this, you will first need to create an account apply for a license.
    \item Inside the software center, find the QNX SDP 8.0 package and install it Once you do, unzip it.
    \item Find the package with the QNX BSP for the Raspberry Pi 4 and install it. Once you do, unzip it.
    \item Find the package that contains QNX Momentics IDE and install it. Once you do, make sure you can open the IDE and import directories as projects. Follow any setup steps it prompts.
\end{enumerate}

\textbf{2) Installing Windows WSL:}
\begin{enumerate}
    \item Open Windows Powershell. Run the command \texttt{wsl --install}.
    \item Open WSL. Navigate to the directory where you installed QNX SDP 8.0. It should be something similar to \texttt{/mnt/c/path-to-qnx-directory/qnx800}.
    \item In the qnx800 directory, set up your work environment with this command: \texttt{source qnxsdp-env.sh}.
\end{enumerate}

\textbf{3) Installing Windows OpenSSH Server:}
\begin{enumerate}
    \item Locate OpenSSH Server and install it.
    \item Open Windows Powershell. Run the following commands to start the SSH server:
    \begin{itemize}
        \item \texttt{Get-Service ssh-agent, sshd | Set-Service-StartupType Automatic}
        \item \texttt{Start-Service ssh-agent}
        \item \texttt{Start-Service sshd}
    \end{itemize}
\end{enumerate}

\textbf{4) Installing PuTTY:}
\begin{enumerate}
    \item Go to the official PuTTY download page, and download the correct msi file for your system.
    \item Locate the msi file and run it to launch the installer. Complete the setup, then make sure you can open the application.
\end{enumerate}

\textbf{5) Installing Liboqs:}
\begin{enumerate}
    \item Go to the QNX Software Center and make sure you have QNX SDP 8.0 BSP for Rasberry Pi BCM2711 R-P14 downloaded
    \item From your local command prompt navigate to you QNX instance and run qnxsdp-env.bat
    \item Install CMake and Ninja
        \begin{verbatim}
        winget install -e --id Kitware.CMake
        winget install -e --id Ninja-build.Ninja
        \end{verbatim}
    \item Build Liboqs
        \begin{verbatim}
            cd C:\
            git clone -b main https://github.com/open-quantum-safe/liboqs.git
            cd liboqs
            rmdir /s /q build-qnx-aarch64

            cmake -S . -B build-qnx-aarch64 -G Ninja ^
              -DCMAKE_SYSTEM_NAME=QNX ^
              -DCMAKE_SYSTEM_PROCESSOR=aarch64 ^
              -DCMAKE_C_COMPILER=qcc ^
              -DCMAKE_CXX_COMPILER=q++ ^
              -DCMAKE_AR:FILEPATH="C:/Users/harly/qnx800_2/host/win64/x86_64/usr/bin/aarch64-unknown-nto-qnx8.0.0-ar" ^
              -DCMAKE_C_FLAGS="-Vgcc_ntoaarch64le" ^
              -DCMAKE_CXX_FLAGS="-Vgcc_ntoaarch64le" ^
              -DCMAKE_INSTALL_PREFIX=C:/liboqs/build-qnx-aarch64/install ^
              -DOQS_BUILD_ONLY_LIB=ON ^
              -DOQS_MINIMAL_BUILD="SIG_ml_dsa_44" ^
              -DBUILD_SHARED_LIBS=OFF ^
              -DOQS_USE_OPENSSL=OFF ^
              DOQS_DIST_BUILD=OFF ^
              -DOQS_OPT_TARGET=generic ^
              -DOQS_USE_ARM_AES_INSTRUCTIONS=OFF ^
              -DOQS_USE_ARM_SHA2_INSTRUCTIONS=OFF ^
              -DOQS_USE_ARM_SHA3_INSTRUCTIONS=OFF ^
              -DOQS_USE_ARM_NEON_INSTRUCTIONS=OFF
            
            cmake --build build-qnx-aarch64
            cmake --install build-qnx-aarch64
        \end{verbatim} 
\end{enumerate}

\subsection{Compile Bootloader (U-Boot)}
Use WSL and clone the U-Boot Git repository (see section 5.1). Once you have it, follow these steps in WSL:
\begin{enumerate}
    \item Navigate to the u-boot directory that you cloned.
    \item Make sure that you have the necessary libraries and that they are up to date. You can do this with \texttt{sudo apt update} followed by \texttt{sudo apt install -y git dosfstools gcc-aarch64-linux-gnu make flex bison bc libssl-dev libgnutls28-dev}.
    \item Before you build the u-boot binary, create a config with the following command: \texttt{CROSS\_COMPILE=aarch64-linux-gnu- make rpi\_4\_defconfig}.
    \item Now compile the u-boot binary with this command: \texttt{CROSS\_COMPILE=aarch64-linux-gnu- make -j\$(nproc)}. This should generate a \texttt{u-boot.bin} file.
    \item Create a directory called \texttt{/boot\_files}. This is where you will store all the necessary files you need to boot the Raspberry Pi. Copy the \texttt{u-boot.bin} file into your \texttt{/boot\_files} directory.
\end{enumerate}

\subsection{Raspberry Pi Firmware Setup}
Use WSL to clone the Raspberry Pi Firmware repository (see section 5.1).
\begin{enumerate}
    \item Navigate to the firmware directory that you cloned, and then into the \texttt{/boot} directory it contains.
    \item Inside the directory there are a number of files. Copy the following files into your \texttt{/boot\_files} directory:
    \begin{itemize}
        \item \texttt{bcm2711-rpi-4-b.dtb}
        \item \texttt{fixup4.dat}
        \item \texttt{fixup4cd.dat}
        \item \texttt{fixup4db.dat}
        \item \texttt{fixup4x.dat}
        \item \texttt{start4.elf}
        \item \texttt{start4cd.elf}
        \item \texttt{start4db.elf}
        \item \texttt{start4x.elf} 
        \item \texttt{rpi4.build}
        \item \texttt{test-ifs.bin}
    \end{itemize}
\end{enumerate}

\subsection{Create Boot Script and Config File}
Next, you are going to want a command that boots the image, as well as a config in order to ensure your boot goes smoothly. In your \texttt{/boot\_files} directory, follow the following steps:
\begin{enumerate}
    \item Create a file called \texttt{boot.cmd}. This will contain a script that will load your image into memory and boot it.
    \item Put the following into your \texttt{boot.cmd} file:
    \begin{verbatim}
echo "Trying to load primary image"

fdt addr -c

# Load the fdt first
setenv fdt_addr ${fdtcontroladdr}

# Load the image into memory
if load mmc 0:1 0x00080000 ${image}; then
    echo "Loaded primary image into memory"
    sleep 5
    go 0x00080000 ${fdt_addr}

# If it could not be loaded try to load a backup
else
    echo "Failed to load primary image memory"
    if test "${image}" = "ifs-rpi4.bin"; then

        # Loading the backup image
        echo "Trying to load backup image"
        if load mmc 0:1 0x00080000 ifs-rpi4.bin.bak; then
            echo "Loaded backup image into memory"
            sleep 5
            go 0x00080000 ${fdt_addr}

        # Indicate that no image could be loaded
        else
            echo "Failed to load backup image"
        fi

    # Specify that the boot image failed
    else
        echo "Boot failed for both primary and backup images"
    fi
fi
    \end{verbatim}
    Here is a quick rundown of what this does:
    \begin{itemize}
        \item The first thing the script does set the device tree address.
        \item The script then loads your image into memory and checks if it was successful. If it was, the image is booted.
        \item If no image was found or it was not loaded correctly, the script tries to find and load a backup image into memory. If this is successful, the script boots the backup image.
    \end{itemize}
    \item The file now needs to be converted into something Linux readable, then into an executable. Run the command \texttt{dos2unix boot.cmd}, and then \texttt{mkimage -A arm -T script -C none -n "Boot Script" -d boot.cmd boot.scr}. This will generate a file called \texttt{boot.scr}.
    \item Create a \textt{config.txt} file. Insert the following into the file:
    \begin{verbatim}
arm_64bit=1
cmdline=startup.txt
device_tree=bcm2711-rpi-4-b.dtb
enable_uart=1
disable_commandline_tags=1
force_turbo=1
gpu_mem=16
max_framebuffers=2
dtparam=sd_poll_once
kernel=u-boot.bin
    \end{verbatim}
\end{enumerate}

\subsection{Create Raspberry Pi Image }
You now need to build the Raspberry Pi image you actually want to run.
\begin{enumerate}
    \item Open your unzipped BSP directory as a new project in QNX Momentics IDE. In the \texttt{/images} directory, find the file \texttt{rpi4.build} file and replace it with \texttt{rpi4.build} the taken from \texttt{/boot\_files}.
    
    \item Make sure the startup line is the following: \texttt{startup-bcm2711-rpi4 -v -D miniuart -W 2500 -u arg -t "below1G:0x11000000:0x20000000"} (this line should be very close to the top of the file).
    \item Under \texttt{/usr/share/terminfo/q/qansi} please insert file path to \newline \texttt{qnx800/target/qnx/usr/share/terminfo/q/qansi} from your local machine. 
    \item Under the SD Memory Card section, ensure that the card is mounted as a readable directory as follows:
    \begin{verbatim}
devb-sdmmc-bcm2711 cam pnp mem name=below1G sdio addr=0xfe340000,irq=158 dos exe=all
waitfor /dev/hd0 10
mount -e /dev/hd0
waitfor /dev/hd0t12 10
mount -t dos /dev/hd0t12 /fs/sd
    \end{verbatim}
    \item You will find the scripts to run OTA updates at the bottom of your build file.
    Here is a brief explanation of how the ML-DSA script works:
    \begin{itemize}
        \item Takes in the new boot image and the ML-DSA signature file
        \item Checks that the image, signature, current image, and embedded ML-DSA public key all exist.
        \item Starts a background sampler using top so CPU and RAM usage can be recorded during the update.
        \item Runs /proc/boot/validate\_mldsa using the embedded public key and the provided signature.
        \item If validation fails: the update is aborted, no install happens, a failure summary is still written to the log.
        \item If validation passes it continues with the following update. 
        \item Creates a backup of the current image as ifs-rpi4.bin.bak
        \item Copies the new image into a temporary staged file, \texttt{ifs-rpi4.bin.new}.
        \item Replaces the active boot image with the staged one, then syncs to flush writes to disk.
        \item Stops the sampler and parses the recorded top output to calculate:total OTA time, average CPU. maximum CPU, average RAM, maximum RAM
        \item Writes those results to ota\_summary.log and raw sample data to ota\_samples.csv.
    \end{itemize}
    \item Finally retrieve the executable files from \texttt{/aarch64\_bin} and place them within the same directory \texttt{/install/archlle/bin}. Then retrieve the \texttt{rsapubkey.pem} and \texttt{mldsa\_pubkey.bin} from the \texttt{signatures} file and similarly put into the directory \texttt{/install/archlle/bin}. 
    \item Now that you are done with \texttt{rpi4.build}, build the project. This should generate a \texttt{ifs-rpi4.bin} image file. Copy this new file into your \texttt{/boot\_files} directory.
\end{enumerate}
\subsection{Sign Raspberry Pi Image}
You now need to create the signature for your bin file
\begin{enumerate}
    \item To create the signature for your bin use the provided files \texttt{scr}: \texttt{sign\_mldsa}, \texttt{keys}: \texttt{mldsa\_prvkey.bin}, and \texttt{Makefile}
    \item using the command make build the executable file for \texttt{sign\_mldsa.c}
    \item Run the following command \texttt{./sign\_mldsa mldsa\_prvkey.bin ifs-rpi4.bin new-ifs.bin.sig}. This will generate your signed bin file
\end{enumerate}

\subsection{Boot the Raspberry Pi}
Your setup should be mostly complete. Now, you need to boot the Raspberry Pi with your QNX image.
\begin{enumerate}
    \item Connect your SD card to your Windows machine. Copy everything from your \texttt{/boot\_files} directory into your SD card. Replace \texttt{ifs-rpi4.bin} with your \texttt{ifs-rpi4.bin} file and a copy called \texttt{new-ifs.bin}. Also add your new-\texttt{ifs.bin.sig} file you created above. Insert your SD card into the Raspberry Pi.
    \item Connect the USB to TTL cable to your windows device, and the following TTL ends to the Raspberry Pi's UART pins:
    \begin{itemize}
        \item The black cable to pin 6
        \item The green cable to pin 8
        \item The white cable to pin 10
        \item DO NOT PLUG IN THE RED CABLE
    \end{itemize}
    \item Open PuTTY, and choose the serial connection type. Check your Windows machine's Device Manager to see what Port is connected (should be similar to COM3). Fill the serial line with the port connection you found, and set the speed to 115200. Press Open, and you should see a blank terminal appear.
    \item Connect the Raspberry Pi to power. The first time you boot, press any key to interrupt the boot, as we want to access the U-Boot terminal. From here, set an environment variable for the image and boot script, and run the script with the following commands:
    \begin{verbatim}
setenv bootcmd 'setenv image ifs-rpi4.bin; fatload mmc 0:1 ${loadaddr}
boot.scr; source ${loadaddr}' (THIS IS ALL ONE LINE)

saveenv

reset
    \end{verbatim}
    \item From here, you should simply be able to watch your image boot. On future boots, you will not need to enter the U-Boot terminal again unless you want to change more environment variables.
\end{enumerate}

\subsection{Run the OTA Update}
Finally, with your image up and running, we can try to run an OTA update.
\begin{enumerate}
    \item A network connection will have to be established first. Plug your Raspberry Pi into your Windows machine with an Ethernet cable.
    \item On the QNX image, set a static IP with the command \texttt{ifconfig genet0 192.168.1.10 netmask 255.255.255.0 up}. Check to see if the network is active with the command \texttt{ifconfig genet0}.
    \item On your Windows machine, open your windows settings. Navigate to Network \& internet, then find the Ethernet connection and click Edit next to IP Assignment, and choose the Manual option. Set the IP Address to 192.168.1.11, the Subnet Mask to 255.255.255.0, and then press save to apply changes.
    \item The two devices should now be able to connect to each other. To test this, use the ping command to see if packets are received from either device. On both the Windows Command Prompt and your QNX OS image, the command syntax should be \texttt{ping <OTHER\_DEVICE\_STATIC\_IP>}. Both should result in packets being received consistently.
    \item On your QNX image, run the following commands to set up the partitions needed for the OTA rollback capabilities.
    \begin{verbatim}
    fdisk /dev/hd0 add -s 2 -t 177 -c $A_START,$A_END
    fdisk /dev/hd0 add -s 3 -t 178 -c $B_START,$B_END
    fdisk /dev/hd0 add -s 4 -t 179 -c $D_START,$D_END

    shutdown
    mount -e /dev/hd0
    ls /dev/hd0t*
    
    mkqnx6fs /dev/hd0t177
    mkqnx6fs /dev/hd0t178
    mkqnx6fs /dev/hd0t179

    mount -t qnx6 /dev/hd0t177 /system
    mount -t qnx6 /dev/hd0t179 /data
    \end{verbatim}
     \item Now to run your ML-DSA validated OTA update enter this command in terminal \texttt{ksh /proc/boot/ota\_update\_ifs\_rsa.sh /fs/boot/new-ifs.bin /fs/boot/new-ifs.bin.sig} 
\end{enumerate}

\section{Results and Analysis}
\subsection{Summary of Findings}
\begin{table}[ht]
\centering
\small
\setlength{\tabcolsep}{4pt} % default is usually 6pt
\caption{Summary of OTA Results}
\begin{tabular}{lcccccccc}
\toprule
  Algorithm & Verify Time & Total OTA & Avg CPU & Max CPU & Avg RAM & Max RAM & Sig Size \\
  & (s) & (s) & (\%) & (\%) & (MB) & (MB) & (bytes) \\
\midrule
Unsigned & N/A         & 6.95\pm 0.17 & 2.73\pm 0.10 & 3.97\pm 0.12 & 201\pm 0       & 201\pm 0 & 0    \\ 
RSA      & 0.8\pm 0.09 & 7.45\pm 0.15 & 3.54\pm 0.05 & 5.27\pm 0.03 & 199.93\pm 0.03 & 200\pm 0 & 256  \\ 
ML-DSA   & 0.7\pm 0.10 & 7.25\pm 0.10 & 3.53\pm 0.05 & 5.33\pm 0.01 & 200.89\pm 0.05 & 201\pm 0 & 2420 \\
\bottomrule
\end{tabular}
\end{table}
\subsection{Signature Verification Time}
Surprisingly, the tests showed a slight reduction in signature verification time between ML-DSA and RSA. This is possibly due to the time taken by OpenSSL's polymorphic key initialization functions to determine which algorithm was used to generate the public key. Overall, however, this did not result in a significant difference in verification time between the two algorithms.
\begin{center}
\begin{tikzpicture}
\begin{axis}[
    title=Figure 2: Signature Verification Time by Algorithm,
    ybar,
    enlargelimits=0.5,
    legend style={at={(0.5,-0.05)},
      anchor=north,legend columns=-1},
    ylabel={Seconds},
    symbolic x coords={},
    xtick=data,
    nodes near coords,
    nodes near coords align={vertical},
    ]
\addplot+ [error bars/.cd, y dir=both,y explicit]
    coordinates {(,0.8) +- (,0.09)};
\addplot+ [error bars/.cd, y dir=both,y explicit]
    coordinates {(,0.7) +- (,0.10)};
\legend{RSA,ML-DSA}
\end{axis}
\end{tikzpicture}
\end{center}
\subsection{Total OTA Time}
The difference in total OTA time was a similar story to section 6.2. ML-DSA surprisingly had a reduction of 0.2 seconds when compared to RSA and was only 0.3 seconds longer than the unsigned option. Again, however, this is not a terribly significant difference.
\begin{center}
\begin{tikzpicture}
\begin{axis}[
    title=Figure 3: Total OTA Time by Algorithm,
    ybar,
    enlargelimits=0.5,
    legend style={at={(0.5,-0.05)},
      anchor=north,legend columns=-1},
    ylabel={Seconds},
    symbolic x coords={},
    xtick=data,
    nodes near coords,
    nodes near coords align={vertical},
    ]
\addplot+ [error bars/.cd, y dir=both,y explicit]
    coordinates {(,6.95) +- (,0.17)};
\addplot+ [error bars/.cd, y dir=both,y explicit]
    coordinates {(,7.45) +- (,0.15)};
\addplot+ [error bars/.cd, y dir=both,y explicit]
    coordinates {(,7.25) +- (,0.10)};
\legend{Unsigned,RSA,ML-DSA}
\end{axis}
\end{tikzpicture}
\end{center}
\subsection{Resource Usage}
CPU usage across the two algorithms did not differ greatly. There was a <= 0.06 percent difference between both their average and max CPU usages. When compared to the unsigned option however this difference jumps to >= 0.8 and >= 1.3 percent for average and max CPU usage respectively. As for RAM usage, all three options remained around 201MB with RSA having a very minor reduction of 1MB.
\begin{figure}[H]
\begin{tikzpicture}
\begin{axis}[
    title=Figure 4: CPU Usage by Algorithm,
    ybar,
    enlargelimits=0.5,
    legend style={at={(0.5,-0.15)},
      anchor=north,legend columns=-1},
    ylabel={Percent},
    symbolic x coords={Average, Max},
    xtick=data,
    nodes near coords,
    nodes near coords align={vertical},
    ]
\addplot+ [error bars/.cd, y dir=both,y explicit]
    coordinates {(Average,2.73) +- (Average,0.10)
                 (Max,3.97) +- (Max,0.12)};
\addplot+ [error bars/.cd, y dir=both,y explicit]
    coordinates {(Average,3.54) +- (Average,0.05)
                 (Max,5.27) +- (Max,0.03)};
\addplot+ [error bars/.cd, y dir=both,y explicit]
    coordinates {(Average,3.53) +- (Average,0.05)
                 (Max,5.33) +- (Max,0.01)};
\legend{Unsigned,RSA,ML-DSA}
\end{axis}
\end{tikzpicture}
\begin{tikzpicture}
\begin{axis}[
    title=Figure 5: RAM Usage by Algorithm,
    ybar,
    enlargelimits=0.5,
    legend style={at={(0.5,-0.15)},
      anchor=north,legend columns=-1},
    ylabel={MB},
    symbolic x coords={Average, Max},
    xtick=data,
    nodes near coords,
    nodes near coords align={vertical},
    ]
\addplot+ [error bars/.cd, y dir=both,y explicit]
    coordinates {(Average,201) +- (Average,0)
                 (Max,201) +- (Max,0)};
\addplot+ [error bars/.cd, y dir=both,y explicit]
    coordinates {(Average,199.93) +- (Average,0.03)
                 (Max,200) +- (Max,0)};
\addplot+ [error bars/.cd, y dir=both,y explicit]
    coordinates {(Average,200.89) +- (Average,0.05)
                 (Max,201) +- (Max,0)};
\legend{Unsigned,RSA,ML-DSA}
\end{axis}
\end{tikzpicture}
\end{figure}

\section{Discussion of Findings}
\subsection{Practical Meaning of the Results}
As seen in section 6, the impact of post-quantum computing signature algorithms on QNX is minimal. Furthermore, when comparing post-quantum solutions such as liboqs to the current OpenSSL standard, these algorithms are actually slightly beneficial to both resource usage and runtime. The biggest drawback comes from the larger key and signature lengths as the added 1056 and 2164 bytes respectively may prove an issue for systems with limited storage. Additionally, for machines which currently choose to forgo implementation of signature verification whatsoever the impacts are much more significant, meaning issues could arise in areas such as CPU usage. This comes mainly in the case of machines making due with the bare minimum spec wise. For most machines, however, it can be concluded that the migration of QNX OS to a post-quantum world is entirely feasible. 
\subsection{Limitations}
\textcolor{red}{Anything about uboot and OTA?}

As of OpenSSL 3.5, the library's polymorphic EVP\_PKEY functions used for all aspects of cryptography support a variety of post-quantum cryptographic algorithms such as SLH-DSA, ML-KEM, and ML-DSA. Unfortunately, however, QNX currently ships with OpenSSL 3.0 meaning that, for the purposes of this paper, we had to use liboqs in its place. This is significant because post-quantum migration will most likely come in the form of QNX updating to this new version of OpenSSL rather than the widespread adoption of liboqs, which is both external and would require large changes to legacy code. Thus, the results of our tests reflect a future different to what is probable.
\subsection{Future Work}
\textcolor{red}{Anything else?}

As mentioned previously in section 7.2, future adoption of post-quantum computing algorithms will most likely come from QNX updating to use OpenSSL 3.5. When this becomes a reality, it is thus imperative to recreate these tests using OpenSSL's built in functionality as opposed to the external liboqs library for accurate results. 
\section{Conclusion}
\Section{References}
\end{document}
